---
  title: "Processing passages - Data Extraction"
output:
  html_document:
  highlight: pygments
theme: journal
toc: yes
toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sprawl)
library(ggplot2)
library(ggjoy)
library(tibble)
library(raster)
library(phenoriceR)
library(dplyr)
library(mapview)
library(RColorBrewer)
library(sf)

mapviewOptions(basemaps = c("Esri.WorldImagery", "OpenStreetMap.DE"),
               raster.palette = colorRampPalette(brewer.pal(9, "RdYlGn")),
               vector.palette = colorRampPalette(brewer.pal(9, "YlGnBu")),
               na.color = "transparent",
               layers.control.pos = "topright")


```

```{r eval = FALSE}
#   ____________________________________________________________________________
#   set input and output folders                                            ####

in_shp <- read_vect("/home/lb/my_data/prasia/Data/vector/Ricetlas/riceatlas_asia.shp")

sub_area  <- "Region 3 - Central Luzon"
in_region <- dplyr::filter(in_shp, ISO == "PHL",
                           REGION == sub_area)
sub_area <- "Reg_3"
in_region <- in_region[1] %>%
 sf::st_combine() %>%
 sf::st_sf(id = 1, sf_column_name = ".") %>%
 sf::st_transform(get_proj4string("/home/lb/my_data/prasia/Data/orig_mosaic/param_series/decirc/eos_decirc.tif"))

mosaic_folder  <- "/home/lb/my_data/prasia/Data/orig_mosaic/param_series/"
subsets_folder <- "/home/lb/my_data/prasia/Data/subsets"
make_folder(subsets_folder, type = "dirname", verbose = T)

pr_extract_subarea(mosaic_folder,
                in_region,
                sub_area,
                out_folder  = subsets_folder)
```


Now, in "out_folder/subsets/subsets_name" we have the "ordered" phenorice tiffs, subsetted
and masked on the region of interest. We can work starting from that folder, now.

Let's see some quick plots:

```{r, message=FALSE, warning=FALSE, eval = TRUE}
subsets_folder <- "/home/lb/my_data/prasia/Data/subsets/"
sosfile <- get(load(file.path(subsets_folder, "Nueva_Ecija/param_series/orig/", "sos.RData")))
boundaries <- get_boundaries("PHL", 1)
# find the bands of "s2" season
which_s2 <- which(regexpr("_s2", names(sosfile)) != -1)
sprawl::plot_rast_gg(sosfile[[which_s2[c(2,4,6,8,10)]]],
title = "SOS doy - Season 2 - Nueva Ecija",
palette_name = "RdYlGn", zlims = c(0.02,0.98), zlims_type = "percs",
outliers_style = "to_minmax", na.color = "transparent",
borders_layer = boundaries,
scalebar_dist = 20)

eosfile <- get(load(file.path(subsets_folder, "Nueva_Ecija/param_series/orig/", "eos.RData")))
NAvalue(eosfile) = -999
sprawl::plot_rast_gg(eosfile[[which_s2[c(2,4,6,8,10)]]],
title = "EOS doy- Season 2 - Nueva Ecija",
palette_name = "RdYlGn", zlims = c(0.02,0.98), zlims_type = "percs",
outliers_style = "to_minmax", na.color = "transparent",
borders_layer = boundaries,
scalebar_dist = 20)

# Extract data on a fishnet and plot as a vector
net <- create_fishnet(sosfile[[1]], cellsize = 3500, shape = "hex")
extr_data <- extract_rast(sosfile[[which_s2]], net, full_data = FALSE)
datain <- extr_data$stats %>%
# dplyr::filter(band_name == "sos_2003_s2") %>%
dplyr::filter(!is.na(avg))
borders <- get_boundaries("PHL", level = 1)
dataplot <- datain %>%
dplyr::filter(band_name %in% c("sos_2003_s2", "sos_2007_s2", "sos_2011_s2",
"sos_2015_s2"))
plot_vect(dataplot, fill_var = "avg",
borders_layer = borders,
palette = "RdYlGn",
na.color = "transparent",
title = "Sowing dates") + facet_wrap(~band_name)

```

## Extract data in an "analysis friendly" format

The format of the sos, eos and pos parameters makes it difficult to further analyze them
because:

1. they use DOY, which is a "circular" variable which is difficult to use in any statistical
analysis. To further add to this problem, in PhenoRice outputs we can have both positive and
negative doys (at least in s1), and doys >365 (at least in s4);
2. data is "artificially" subdivided in the 4 "seasons" specified in PhenoRice processing.
This subdivision doesn't make any sense from an "agronomic" point of view: we need to be
able to "collapse" all data for a given area into a **single time series of event occurrence**
  (sos, pos or eos) covering the 14 years of analysis, on which we can later make statistical analysis.

To do so:

  ### "Decircularize" the doy-year information

  By "decircularize" I mean going from the DOY representation to a Julian Date representation,
which has the advantage of being a continouos monotonically increasing variable.

sos, eos and pos dates are therefore here converted to a numeric variable representing the
number of days elapsed since 01/01/2000. (Note however that it is always possible to go back
                                          to the date by simply "adding" the values reported to the raster to 01/01/2000)

(This passage is very slow on the full mosaics because I didn't have time yet to optimize
  it, but since we need to do it only once we can live with it for the mkoment. Moreover, when
  working on smaller subsets it is reasonably fast, so I'll stay like this at the moment)

```{r eval = FALSE, message=FALSE, warning=FALSE, eval = FALSE}
in_folder  <- "/home/lb/my_data/prasia/Data/orig_mosaic/param_series/orig"
out_folder <- file.path(in_folder, "decirc")
decirc_phenorice(in_folder, out_folder)
```

Let's mask again the de-circularized data on Nueva Ecija:

```{r, message=FALSE, warning=FALSE, eval = TRUE}
in_folder <- "/home/lb/my_data/prasia/Data/orig_mosaic/param_series/decirc"
r <- read_rast("/home/lb/my_data/prasia/Data/orig_mosaic/param_series/decirc/sos_decirc.tif")
subset_name <- "Nueva_Ecija"
in_country  <- "PHL"
boundmask   <- sprawl::get_boundaries(in_country, level = 1) %>%
sf::st_as_sf() %>%
dplyr::filter(NAME_1 == "Nueva Ecija") %>%
sf::st_as_sf()%>%
sf::st_transform(get_proj4string("/home/lb/my_data/prasia/Data/orig_mosaic/param_series/orig/cumevi_ordered.tif"))


# insos_decirc_NE <- mask_rast(r, boundmask, crop = T,
#                              out_filename = file.path(in_folder,"subsets/NE/insos_NE.tif"),
#                              parallel = T)
insos_decirc_NE <- read_rast(file.path(in_folder,"subsets/NE/insos_NE.tif"))
NAvalue(insos_decirc_NE) <- 0
summary(insos_decirc_NE[[1:8]])
```

You see that the "dates" are more or less monotonically increasing on the different seasons,
though there are some overlaps. Analyzing the seasons separately would therefore be
difficult. So, it's better to put everyting in a single array (note that here I can do this brute
                                                               force, but for larger areas it will be impossible):

  ```{r, message=FALSE, warning=FALSE}
# get the data for three years as an array that contains all detected sos dates in Nueva
# Ecijia as a single array
insos_decirc <- get(load(
  "/home/lb/my_data/prasia/Data/subsets/Nueva_Ecija/param_series/decirc/sos.RData"
))
sos = (as.Date("2000-01-01") + raster::getValues(insos_decirc[[1:24]]))
sos = sos[!is.na(sos)]
alldata = data.frame(sos = sos) %>%
  as_tibble()

ggplot(alldata, aes(x = sos)) +
  geom_histogram(binwidth = 8) +
  scale_x_date(date_breaks = "1 year", limits = as.Date(c("2002-06-01", "2008-12-01"))) +
  theme_bw() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

This clearly depicts the double seasonality of Nueva Ecjia. It also evidences the rather
narrower range of detected sos in winter.

However, what I wanted to point out is that "decircularizing" the DOYS allows to analyze
them more easily, becaus we can "treat" the detected sos as a whole time series of events,
without having to think about to which "season" PhenoRice assigned the sos.

### Aggregate to lower resolution

To be able to work with a manageable number of "pixels" over large areas, and also to create
large-scale maps, it could be worthy to aggregate the results over regular grids:

  ```{r, message=FALSE, warning=FALSE}

in_fishnet <- sprawl::create_fishnet(insos_decirc, pix_for_cell = 21,
                                     shape = "hex")
plot_rast(insos_decirc[[2]], in_poly = in_fishnet)
insos_decirc_5k <- extract_rast(insos_decirc,
                                in_fishnet,
                                join_geom = FALSE,
                                verbose = FALSE)

insos5kdata <- insos_decirc_5k$alldata %>%
  dplyr::filter(!is.na(value))  %>%
  dplyr::mutate(sosdate = as.Date("2000-01-01") + value,
                fc = N/n_pix,
                area = 2146.58*fc) %>%
  dplyr::arrange(id_feat, band_n)
insos5kdata

summary(insos5kdata$fc)
```

here we get, for each 5km cell, the average sos date estimated in the different years and
seasons, the number of pixels detected within the cell from which we can compute the
detected area (so that we can filter-out 5k cells where only few pixels where rice - in this
               case you can see that half of the 5k cells got a 50 % rice "detection").

Again, having de-circulated the doys allows to treat all data as a continuous series.
Here I plot the frequency histograms derived by considering all Nueva Ecjia cells with fc
above 0.5:

  ```{r, message=FALSE, warning=FALSE, eval = TRUE}
insos5kdata_filt <- insos5kdata %>%
  dplyr::filter(fc >= 0.5)
insos5kdata_filt

# plot the distibutioni of detected dates in cells above 50%

ggplot(insos5kdata_filt, aes(x = sosdate)) +
  geom_histogram(binwidth = 8) +
  scale_x_date(date_breaks = "1 year", limits = as.Date(c("2003-06-01", "2016-12-01"))) +
  theme_bw() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

To analyze interannual variations, we will have however to "split" the series somewhere,
but now we can decide where. Here I try with october.

```{r, message=FALSE, warning=FALSE, eval = TRUE}
years <- 2002:2016
season <- cut((insos5kdata_filt$sosdate),
              breaks = as.Date(paste(years,"-01-10",sep="")),
              labels = paste(years[-length(years)],years[-length(years)]+1,sep="/"))

insos5kdata_filt <- insos5kdata_filt %>%
  mutate(date_fake = doytodate(datetodoy(sosdate), 2003)) %>%
  mutate(season = season) %>%
  dplyr::filter(!season %in% c("2002/2003", "2001/2002"))

ggplot(insos5kdata_filt, aes(x = date_fake, y = season, fill = season)) +
  geom_joy(scale = 2.0) +
  scale_x_date(limits = as.Date(c("2002-11-01", "2003-11-01")),
               date_labels = "%b %d") + theme_joy() +
  scale_fill_cyclical(values = c("gray80", "gray50")) +
  ggtitle("Nueva Ecija - Interannual variation of detected SOS")
```

(This is just a first try: I probably have some error in the grouping, but to give you an
 idea)

That's all for now. I'll continue working on this in the next days. In particular, I plan
to start working on the Rice Atlas, and also on functions to analyze temporal trends (for
                                                                                      that, I think we'll have to define some methods for clustering the data, or at least identify
the main modes in the distribution)
