---
title: "Processing passages - Data Extraction Simplified"
output:
  html_document:
  highlight: pygments
theme: journal
toc: yes
toc_depth: 2
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tibble)
options(tibble.print_max = Inf) 
options(dplyr.width = Inf)
```

This page documents the passages required to extract PhenoRice results for the 
different polygons of the RiceAtlas dataset. Input data are the raster mosaics
stored in `Data/orig_mosaics`, and the reshuffled riceAtlas shapefile stored 
in `Data/vector/Ricetlas/riceatlas_asia_reshuffled.shp`

# Extract data for a specific RiceAtlas region

Suppose you want to extract data for Region: "Region_3_-_Central_Luzon":  

```{r  message=FALSE, warning=FALSE, eval=FALSE}
#   ____________________________________________________________________________
#   set input and output folders                                            ####

# set the main folder ----
main_folder <- "/home/lb/my_data/prasia/Data"

#'  # Suppose you want to extract data for Region: "Region_3_-_Central_Luzon" :
#'  # --> extract it from the full shapefile

Region_name <- "Region_3_-_Central_Luzon"

## "1) run "pr_extract" on the region ----

pr_extract(main_folder,
           region = Region_name)

```

__NOTE:__ Using `region = "All"` will automatically process all RiceAtlas regions. 

The function crops the original mosaics on the selected region, and then extracts
the data for the different sub-regions of the specified RiceAtlas regions. 

- Cropped rasters are stored in the `Data/subsets/_region_name_/orig` and
`Data/subsets/_region_name_/decirc` folders. 

- Extracted data are saved in the `Data/subsets/_region_name_/RData` folder. One
separate RData file is created for each PhenoRice parameter (e.g., 
"Dhaka_sos_stats.RData", "Dhaka_eos_stats.RData", "Dhaka_sos_decirc_stats.RData")

# Accessing the extracted data

Extracted data can be successively accessed by loading the saved RData files. Here I show it 
for the "sos" dataset of "Central_luzon". 

```{r}

main_folder <- "/home/lb/my_data/prasia/Data"
indata <- get(load(file.path(main_folder,
                             "subsets/Region_3_-_Central_Luzon/RData/Region_3_-_Central_Luzon_sos_stats.RData")))

```
The data is saved as a list containing two elements:

1) **`indata$alldata`** contains "raw" info extracted __from all pixels__ of the selected
region that were recognized as "rice" in the different years: 

```{r}

head(indata$alldata)

```
It contains the following columns: 

-  __ID_name__  : Name of RiceAtlas polygon 
-  __variable__ : PhenoRice variable      
-  __band_n__   : band of the tiff mosaics from which values come from
(e.g., 1 = 2003 seas. 1; 2 = 2003 seas 2....)       
-  __year__     : year of the analysis (note that "date" can correspond to a 
different year on seasons spanning end of the yea (e.g., year = 2003, sos date = 2002-10-25)       
-  __date__     : date of occurrence of the PhenoRice metric
-  __doy__      : doy of occurrence of the PhenoRice metric (for the "decirc" variables
this is the number of days since 2000-01-01; for "standard" varibales, it is 
number of days since the beginning of the year (negative values possible as well 
as values above 365!))
-  __N__        : unique identifier of a pixel within the subregion
-  __n_pix_val__: number of pixels identified as being rice within the subregion


so, for example, subsetting on a specific "ID_NAME" gives us
all PhenoRice results for a given subregion: 

```{r}
require(dplyr)
pix_data <- indata$alldata %>% 
  dplyr::filter(ID_name == "PHL_Region 3 - Central Luzon_Zambales")
head(pix_data)
```

from which we will be able to extract all kind of information without bothering 
anymore with the homungus raster mosaics! For example: 

```{r}
library(ggplot2)
ggplot(pix_data) + geom_violin(aes(x = factor(year), y = doy), alpha = 0.4) + 
  theme_bw() + 
  ggtitle("Sos Dates distribution - Central Luzon_Zambales") + 
  coord_flip()

```

2) **`indata$stats`** contains "summarized" data over each subregion and PhenoRIce
season: 

```{r message=FALSE, warning=FALSE}

head(indata$stats)

```

-  __ID_name__  : Name of RiceAtlas polygon 
-  __variable__ : PhenoRice variable      
-  __band_n__   : band of the tiff mosaics from which values come from
(e.g., 1 = 2003 seas. 1; 2 = 2003 seas 2....)       
-  __year__     : year of the analysis (note that "date" can correspond to a 
different year on seasons spanning end of the yea (e.g., year = 2003, sos date = 2002-10-25)       
-  __avgdate__  : Average date of occurrence of the PhenoRice metric __over the 
"band" and sub region__
-  __avgdoy__   : Average doy of occurrence of the PhenoRice metric __over the 
"band" and sub region__ (the same conventions for decirc and original variables
described above apply)
-  __meddoy__   : Median doy of occurrence of the PhenoRice metric __over the 
"band" and sub region__
-  __sd__       : Standard deviation of doy of occurrence of the PhenoRice metric 
__over the "band" and sub region__
-  __mindoy__   : Minimum doy of occurrence of the PhenoRice metric __over the 
"band" and sub region__
-  __maximum__   : Maximum doy of occurrence of the PhenoRice metric __over the 
"band" and sub region__  
-  __n_pix_val__ : number of pixels identified as being rice within the subregion
-  __n_pix__     : total number of pixels within the subregion


(Note that for the current analysis this is not very useful, since the summaruized
data are computed on a "per phenorice season" basis. It is better to use the "alldata"
element of the output and perform any needed data extraction / summarization starting
from that.)

# Next steps

### __Preliminary analysis of consistency between PhenoRice and RiceAtlas__

by comparing
Phenorice histograms with riceatlas data. Histograms can be quickly produced like
this:

```{r message=FALSE, warning=FALSE}
library(sprawl)
# Create a dummy variable, aggregating all data as if they were from the same year
indata$alldata$fakedate <- sprawl::doytodate(indata$alldata$doy, 2003)

library(ggplot2)
ggplot(indata$alldata) +
  geom_histogram(aes(x = fakedate), binwidth = 8) +
  facet_wrap(~ID_name, scales = "free_y") + theme_bw()
```

Info about the riceatlas season area and the sos ranges __should be overplotted
for reference to allow a quick comparison__. 

### __mixtools analysis to "identify" the main seasons__ 

Needs to be done for each RiceAtlas polygon. We should decide if:  

a) Subsetting the data on a selected year, and analyse only that one. 
b) Analysing each year separately 
c) Aggregating all data as if they were from the same year (this allows to 
increase the dimensions of the sample in low rice fc areas). This requires
creating a "dummy variable" using something like: 

```{r message=FALSE, warning=FALSE}

indata$alldata$fakedate <- doytodate(datetodoy(indata$alldata$date), 2003)
head(indata$alldata)
```

### __Compute the PhenoRice Areas__

Once the main seasons are identified, __compute the area__ detected by PhenoRice
within each season (by subsetting on date/fakedate)

### __Summarize the Comparison__

Find a "statistical" way to __compare both the detected rice seasons and areas 
  and summarize results__ of the comparison.



