c("plant_st1   105   181   258       4323")
title: "RiceAtlas Comparison"
output:
html_document:
highlight: pygments
theme: journal
toc: yes
toc_depth: 2
editor_options: 
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sprawl)
library(ggplot2)
library(ggjoy)
library(tibble)
library(raster)
library(phenoriceR)
library(dplyr)
library(mapview)
library(RColorBrewer)

mapviewOptions(basemaps = c("Esri.WorldImagery", "OpenStreetMap.DE"),
               raster.palette = colorRampPalette(brewer.pal(9, "RdYlGn")),
               vector.palette = colorRampPalette(brewer.pal(9, "YlGnBu")),
               na.color = "transparent",
               layers.control.pos = "topright")


```

Here I am going to document the passages required to compare PhenoRice results
with RiceAtlas

# Preprocessing

## Select an area

For simlpicity, let's work on a selected area (in this case, the Philippines) 
(for Bhogendra: I created a folder structure containing all required input 
data. It would be good if you could replicate it on your machine so that we
can more easily share the code)

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# This is the main folder containing preprocessed PhenoRice global mosaics
# in the form of multiband TIFF images for each parameter (in "/orig"), as 
# well as the results of "decircularization (in "/decirc")
mosaic_folder  <- "/home/lb/my_data/prasia/Data/orig_mosaic/param_series/"
subsets_folder <- "/home/lb/my_data/prasia/Data/subsets"
make_folder(subsets_folder, type = "dirname", verbose = T)
subset_name <- "PHL"
in_country  <- "PHL"
boundmask   <- sprawl::get_boundaries(in_country, level = 1) %>%
  sf::st_as_sf() %>% 
  sf::st_combine() %>%
  sf::st_sf(id = 1, sf_column_name = ".") 
plot_vect(boundmask, scalebar_dist = 200, show_axis = TRUE)

# Extract PhenoRice data for the PHL subset area - cropped and masked results are 
# placed in "prasia/Data/Subsets/PHL"

extract_subarea(mosaic_folder,
                boundmask,
                subset_name,
                out_folder  = subsets_folder)

#Extract RiceAtlas data for PHL

```

## Visualize data

Extract and visualize rice atlas data for an area and "variable -  in this case
"Peak" planting date. 

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
in_txt <- "/home/lb/my_data/prasia/various/NormalizedDB.txt"
in_riceatlas_shp = "/home/lb/my_data/prasia/Data/vector/Ricetlas/riceatlas_asia.shp"
riceatlas_phl <- read_riceatlas(in_txt, in_riceatlas_shp, "PHL")
peak_plant_s1 <- subset(riceatlas_phl, Pheno_stage == "plant_st1")
plot_vect(peak_plant_s1, fill_var = "Peak", palette_name = "RdYlGn")
```

Here we immediately __see something strange__. The "dark green" polygon has
completely different values. Let's see a histogram:  

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
p <- ggplot(peak_plant_s1, aes(x = doytodate(Peak, 2005))) + 
  geom_histogram(binwidth = 15, fill = "darkgrey", color = "black") +
  scale_x_date(limits = doytodate(c(0,365), 2005) ,
               date_labels = "%b") + 
  theme_bw() 
p

```
Looking at the original table of the `rice_production_v1` shapefile, this is
confirmed (i.e., it is not a data conversion problem). __The "correct" dates for 
the Wet Season of the offending polygons is instead reported in this case in 
"plant_st_2"__:

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data_leyte <- subset(riceatlas_phl, Sub_region %in% c("Leyte" , "Samar") & 
                       Pheno_stage %in% c("plant_st1", "plant_st2")) %>% 
  select(Sub_region, Season, Peak) 
st_geometry(data_leyte) = NULL
data_leyte$Peak <- doytodate(data_leyte$Peak, year = 2005) %>% 
  format("%d - %b")
knitr::kable(data_leyte)
```

You can see that __the two seasons are "reversed" in Leyte__. The problem is that 
if these kind of errors are "widespread" among the DB, then making the comparisons
with PhenoRice may be difficult. 

Additionally, __comparing the `rice_production_v1` and `rice_calendar` data show 
inconsistencies__. For example, in `rice_calendar`, South Leyte polygon has
correct data, while in `rice_production_v1` they are wrong. However, though 
in general the other polygons have similar values, they are not identical. 

Finally, another problem is that, as in original PhenoRice outputs, __the DOYs are 
not "decircularized"__. This makes things difficult because area with the same 
calendar can have for example "peak date" for planting equal to 360 or 10. 
I will have to think of a way to solve this problem, if we wish to compare 
PhenoRice with RiceAtlas. 

### Update after Bhogendra reprocessing of riceAtlas data

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
in_txt <- "/home/lb/my_data/prasia/various/NormalizedDB_Asia_new.txt"
in_riceatlas_shp = "/home/lb/my_data/prasia/Data/vector/Ricetlas/riceatlas_asia.shp"
riceatlas_phl <- read_riceatlas(in_txt, in_riceatlas_shp, "PHL")
peak_plant_s1 <- subset(riceatlas_phl, Pheno_stage == "plant_st1")
plot_vect(peak_plant_s1, fill_var = "Peak", palette_name = "RdYlGn")
```

``` {r}

riceatlas_phl %>% 
  select(Sub_region, N_seasons, Season, Pheno_stage, Start, Peak, End)

```
Problem is evidently related to not accounting for circular nature of DOY. 

### Further reshaping of riceatlas data

Reshape a bit the dataset: 

```{r reshape_data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Reshape a bit the dataset
in_atlas <- read.csv2(in_txt) %>% 
  # split columns
  tidyr::separate(Pheno_stage, c("Stage", "season"), "_st") %>% 
  # remove empty rows
  dplyr::filter((Start != 0 & Peak != 0)) %>% 
  # create unique name
  dplyr::mutate(reg_name = paste(ISO, Region, Sub_region)) %>% 
  dplyr::mutate(reg_name = stringr::str_replace(reg_name, "NA", "")) %>% 
  # Rename and reorder
  dplyr::select(Objectid, ISO, reg_name, N_seasons, season, 
                Stage, Start, Peak, End, Total_area, Seasonal_area, 
                Seasonal_production, Season, Country, Region, Sub_region) %>% 
  dplyr::rename(cty = ISO, N_seas = N_seasons, Seas_name = Season, Season = season,
                Seas_area = Seasonal_area , Seas_prod = Seasonal_production, 
                Tot_area = Total_area) %>% 
  tibble::as_tibble()
names(in_atlas)[1] = "OBJECTID"
in_atlas
```

Now let's compute the differences in the Start/Peak/End columns and check their 
distribution: 

```{r}

in_atlas <- in_atlas %>% 
  dplyr::mutate(endstart_lgt = (End - Start), 
                peakstart_lgt = (Peak - Start)) %>% 
  dplyr::select(1:9, endstart_lgt, peakstart_lgt, everything())

p <- ggplot(in_atlas) + 
  geom_jitter(aes(x = Season, y = endstart_lgt), size = 0.4, alpha = 0.5) + 
  facet_wrap(~Stage) + theme_bw()
p + ggtitle("Distribution of the differences between End and Start in riceatlas")

p <- ggplot(in_atlas) + 
  geom_jitter(aes(x = Season, y = peakstart_lgt), size = 0.4, alpha = 0.5) + 
  facet_wrap(~Stage) + theme_bw()
p + ggtitle("Distribution of the differences between Peak and Start in riceatlas")
```

From this, it is obvious that the dataset is generally coherent, but also that:

1. In some cases the Start/Peak/End values do not account for "circularity". Therefore we can have cases such as:
- `Start = 300 ,Peak = 350, End = 15` --> `End <=Start` and `Peak <= Start` --> leading to negative outliers
- `Start = 15, Peak = 350, End = 320` --> `End <=Start` and `Peak <= Start` --> leading to positive outliers

2. 

riceatlas_phl2 <- riceatlas_phl %>% 
# split columns
tidyr::separate(Pheno_stage, c("stage", "season"), "_st") %>% 
# remove NA rows
dplyr::filter((Start != 0 & Peak != 0))  %>% 
# remove NA rows
dplyr::filter((Start != 0 & Peak != 0))  %>% 
# Compute start vs end and start vs peak differences
dplyr::mutate(diff_end_start = (End - Start), 
diff_peak_start = (Peak - Start))


```

