---
title: 'Validation of Phenorice rice map - Italy'
output:
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    keep_tex: yes
  geometry: margin = 0.5 cm
---


```{r load_data, echo=FALSE, message = FALSE, error = FALSE, hide = TRUE, results = 'hide', warning = FALSE}
 
#---- Load libraries and load data from results of LB_Phenorice_plot_dates.R ----

  library(ggplot2)
  library(reshape)
  library(data.table)
  library(rgdal)
  library(SDMTools)
  library(plyr)
  library(gdalUtils)
  library(raster)
  library(foreign)
  library(tools)
  library(hydroGOF)
  library(gridExtra)
  library(scales)
  library(latticeExtra)


  main_folder = 'D:/Temp/PhenoRice/Processing/PHL/Outputs/2013_new_WS/' 															# Folder used to store results of " 
  results_folder = file.path(main_folder,'Validate')			# Where to put results
  country_code = 'PHL'
  sel_quart = c()
  out_RData_file = file.path(results_folder, 'Plots_Validate_quart3.RData')
  load(out_RData_file)

```
# Introduction 

This report summarize results of PhenoRice rice area validation on the Italian Study area. The analysis is relative to the 2012-2013 growing seasons.
Phenorice was run in this case with "standard" parameterization (4 equal sized seasons - default values for criteria), __,  considering the EVI decrease criteria__

\pagebreak

# Accuracies computed on areas 
```{r accuracies, echo=FALSE, fig.height = 7, fig.cap = 'MODIS vs SIARL area + User s and producers accuracy for rice class - computed on Areas', hide = TRUE, message = FALSE, error = FALSE, results = 'hide' }
  grid.arrange(p_area_comp_all, p_accuracy, p_accuracy_pro)
```

\pagebreak


# Detection Rate

```{r detection, echo=FALSE, fig.height = 8, fig.cap = 'Detection rate vs rice surface in MODIS pixel', hide = TRUE, message = FALSE, error = FALSE, results = 'hide' }
  
  p_detection

```

\pagebreak

# Scatterplots

Comparison between MODIS and Reference rice areas, based on aggregation on 2x2 , 5x5 etc. grids

```{r scatter, echo=FALSE, fig.height = 9, fig.cap = 'Comparison of areas', hide = TRUE, message = FALSE, error = FALSE, results = 'hide' }
  
  grid.arrange(plots_accuracy_1,plots_accuracy_2,plots_accuracy_3,plots_accuracy_4)

```

\pagebreak

# Maps Comparison

Comparison between MODIS and Reference rice maps, based on aggregation on 2x2 , 5x5 etc. grids

```{r maps1, echo=FALSE, fig.height = 9, fig.cap = 'Comparison of MODIS and reference map - 2x2 km grids', hide = TRUE, message = FALSE, error = FALSE, results = 'hide' }
  a = maps_sar[[1]]
  b = maps_modis[[1]]
  grid.arrange(a,b)

```

\pagebreak


```{r maps2, echo=FALSE, fig.height = 9, fig.cap = 'Comparison of MODIS and reference map - 5x5 km grids', hide = TRUE, message = FALSE, error = FALSE, results = 'hide' }
  a = maps_sar[[2]]
  b = maps_modis[[2]]
  grid.arrange(a,b)
```

\pagebreak


```{r maps3, echo=FALSE, fig.height = 9, fig.cap = 'Comparison of MODIS and reference map - 10x10 km grids', hide = TRUE, message = FALSE, error = FALSE, results = 'hide' }
  a = maps_sar[[3]]
  b = maps_modis[[3]]
  grid.arrange(a,b)
```

\pagebreak

s
```{r maps4, echo=FALSE, fig.height = 9, fig.cap = 'Comparison of MODIS and reference map - 20x20 km grids', hide = TRUE, message = FALSE, error = FALSE, results = 'hide' }
  a = maps_sar[[4]]
  b = maps_modis[[4]]
  grid.arrange(a,b)
```
