---
title: "RiceAtlas: Checks and Reshuffling of data"
output:
html_document:
highlight: pygments
theme: journal
toc: yes
toc_depth: 2
editor_options:
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sprawl)
library(ggplot2)
library(ggjoy)
library(tibble)
library(raster)
library(phenoriceR)
library(dplyr)
library(mapview)
library(RColorBrewer)

mapviewOptions(basemaps = c("Esri.WorldImagery", "OpenStreetMap.DE"),
               raster.palette = colorRampPalette(brewer.pal(9, "RdYlGn")),
               vector.palette = colorRampPalette(brewer.pal(9, "YlGnBu")),
               na.color = "transparent",
               layers.control.pos = "topright")
```
Here I am going to document the passages used to reshuffle the RiceAtlas dataset, 
and correct inconsistences, needed to allow an easier comparison with PhenoRice
results.

# Preprocessing - changes on data organization 

The original "RiceAtlas" attribute table is organized in "wide" format, with 
one row for "subregion", and many columns containing the data for the 
different seasons (e.g., __PLANT_ST_1__, __PLANT_ST_2__, ....). 

```{r echo=TRUE, message=FALSE, warning=FALSE}

in_riceatlas_shp = read_vect("/home/lb/my_data/prasia/Data/vector/Ricetlas/riceatlas_asia.shp") %>% 
  arrange(COUNTRY, OBJECTID)
as.data.frame(head(in_riceatlas_shp, 2))
```

This makes working with the dataset difficult. First thing we (Bhogendra) did, was 
transforming the dataset in a more "manageable" long format: 

```{r echo=TRUE, message=FALSE, warning=FALSE}

in_data <- data.table::fread("/home/lb/my_data/prasia/various/NormalizedDB.txt")
head(in_data)
```

Now, we have multiple rows per subregion, each corresponding to a possible "Season"
and to a Phenological "indicator" (e.g., Planting or Harvesting). We can do some additional
reshuffling for ease of use: 

```{r echo=TRUE, message=FALSE, warning=FALSE}

in_atlas <- in_data %>%
  # split the "Pheno_stage" column
  tidyr::separate(Pheno_stage, c("Stage", "season"), "_st") %>%
  # remove empty rows
  dplyr::filter((Start != 0 | Peak != 0 | End != 0)) %>%
  # create a unique name for each region
  dplyr::mutate(reg_name = paste(ISO, Region, Sub_region)) %>%
  dplyr::mutate(reg_name = stringr::str_replace(reg_name, "NA", "")) %>%
  # Rename and reorder columns
  dplyr::select(Objectid, ISO, reg_name, N_seasons, season,
                Stage, Start, Peak, End, Total_area, Seasonal_area,
                Seasonal_production, Season, Country, Region, Sub_region) %>%
  dplyr::rename(cty = ISO, N_seas = N_seasons, Seas_name = Season, Season = season,
                Seas_area = Seasonal_area , Seas_prod = Seasonal_production,
                Tot_area = Total_area) %>%
  dplyr::mutate(Stage = factor(Stage,
                               levels = c("plant", "harv"),
                               labels = c("Planting","Harvesting"))) %>%
  tibble::as_tibble()
names(in_atlas)[1] = "OBJECTID"
head(in_atlas)
```

# Checks on "consistency of the dataset"

A quick check on the "consistency" of the dataset can be made by computing the 
differences between the differences beteween the reported "Start", "Peak" and 
"End" DOYs, for each region. 
```{r}

in_atlas <- in_atlas %>% 
  dplyr::mutate(endstart_lgt  = (End - Start), 
                peakstart_lgt = (Peak - Start),
                endpeak_lgt   = (End - Peak)) %>% 
  dplyr::select(1:9, endstart_lgt, peakstart_lgt, endpeak_lgt, everything())
head(in_atlas)
```

```{r echo=FALSE, fig.height=3, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}

in_atlas <- in_atlas %>% 
  dplyr::mutate(endstart_lgt  = (End - Start), 
                peakstart_lgt = (Peak - Start),
                endpeak_lgt   = (End - Peak)) %>% 
  dplyr::select(1:9, endstart_lgt, peakstart_lgt, endpeak_lgt, everything())

print(ggplot(in_atlas) + 
        geom_histogram(aes(x = endstart_lgt), binwidth = 7)  + 
        facet_wrap(~Stage) + theme_bw() +
        ggtitle("Differences between End and Start in riceatlas"))

print(ggplot(in_atlas) + 
        geom_histogram(aes(x = peakstart_lgt), binwidth = 7) + 
        facet_wrap(~Stage) + theme_bw() + 
        ggtitle("Differences between Peak and Start in riceatlas"))

print(ggplot(in_atlas) + 
        geom_histogram(aes(x = endpeak_lgt), binwidth = 7) + 
        facet_wrap(~Stage) + theme_bw() + 
        ggtitle("Differences between End and Peak in riceatlas"))


```

From this, it is obvious that the dataset is usually coherent, but also that it has 
some problems. In several cases Start/Peak/End values do not account for "circularity".

Therefore we can have cases such as:
- `Start = 354 ,Peak = 3, End = 18` --> `End <=Start` and `Peak <= Start` --> leading to negative outliers
- `Start = 4, Peak = 292, End = 34` --> `End <=Start` and `Peak <= Start` --> leading to positive outliers

, plus other less common problems (e.g., `Start = 32 ,Peak = 32, End = 0`,
`Start = 148, Peak = 148, End = 155`
`Start = 213 ,Peak = 227, End = 213`, `Start = 277, Peak = 19, End = 307`), as well
as obvious mistakes (e.g., `Start = 74, Peak = 74, End = 196`, `Start = 1, Peak = 4, End = 66`)

(It also shows a very anomalous high frequency at `Peak - Start = 30`/ 
`End - Peak = 15`...)

All this makes an automatic comparison with PhenoRice results difficult. For example, 
see what we have as `Start - Peak - End` ranges for some provinces in PHL: 

```{r fig.width=7}
in_phl <- in_atlas %>% 
  dplyr::filter(cty == "PHL" & Region == "Region 3 - Central Luzon")

ggplot(data = in_phl, aes(x = Sub_region)) +
  geom_point(aes(y = Peak, color = Season), position = position_dodge(.3)) + 
  geom_errorbar(aes(ymax = End, ymin = Start, color = Season), 
                position = position_dodge(width = .3)) +
  facet_wrap(~Stage) + coord_flip() + theme_bw() + ylab("DOY") + xlab("RiceAtlas polygon")


```

A mix of correct data, some missing data (e.g, Tarlac Harvesting Start), and 
inconsistent data (e.g., Nueva Ecjia Planting; Bataan Harvesting). Most of the 
problems are obviously related to "circularization" of DOYS. 

# Reshufflling of RiceAtlas data to solve the problem

To solve the problem, we need to filter out errors and reshuffle the DOYS
in a consistent way. 

```{r fig.height=3, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}

out_atlas <- in_atlas %>% 
  # add 365 to End/Peak where differences are negative
  lbscripts::mutate_when((endstart_lgt < 0 & endstart_lgt < 0), 
                         list(End = End + 365)) %>%
  # re-create the diffs for test purposes
  dplyr::mutate(endstart_lgt = (End - Start),
                peakstart_lgt = (Peak - Start), 
                endpeak_lgt   = (End - Peak)) %>% 
  # Second fix where we still have prblems
  lbscripts::mutate_when((endpeak_lgt > 250 & peakstart_lgt < -250 & !is.na(Start)), 
                         list(Start = Start - 365)) %>% 
  lbscripts::mutate_when((endpeak_lgt > 250 & peakstart_lgt < -250 & !is.na(End)), 
                         list(End   = End - 365)) %>% 
  # re-create the diffs for test purposes
  dplyr::mutate(endstart_lgt = (End - Start),
                peakstart_lgt = (Peak - Start), 
                endpeak_lgt   = (End - Peak)) %>% 
  # Third fix where we still have prblems
  lbscripts::mutate_when(endpeak_lgt > 300, 
                         list(End   = End - 365)) %>% 
  # re-create the diffs for test purposes
  dplyr::mutate(endstart_lgt = (End - Start),
                peakstart_lgt = (Peak - Start), 
                endpeak_lgt   = (End - Peak)) %>%
  # Remove Start/End when they are equal to Peak
  lbscripts::mutate_when((Start == Peak), list(Start = NA)) %>%
  lbscripts::mutate_when((End == Peak),   list(End   = NA)) 

# Now manually fixErros and corner cases
out_atlas <- out_atlas %>% 
  #manually fix corner cases
  #"IND Puducherry Yanam", which has very strange data on the original table
  #( Harvesting:  Start  = 4  Peak = 292  End = 34;  Planting: Start  = 277   Peak = 19   End = 307).
  #The obnly reasonable way IMO is to remove Start and set End to 365 + 34 (Planting)
  #and remove Start and End (Harvesting)
  lbscripts::mutate_when((reg_name == "IND Puducherry Yanam" & Stage == "Harvesting" & Season == 2),
                         list(Start = NA, 
                              End = 365 + End)) %>%
  lbscripts::mutate_when((reg_name == "IND Puducherry Yanam" & Stage == "Planting" & Season == 2),
                         list(Start = NA, 
                              End = NA)) %>% 
  #"MMR Tanintharyi"
  lbscripts::mutate_when((reg_name == "MMR Tanintharyi " & Stage == "Planting" & Season == 2),
                         list(Start = NA, 
                              End   = NA)) %>% 
  #"severla places with "213   227   213"" --> Remove End
  lbscripts::mutate_when((Start == 213 & Peak == 227 & End == 213),
                         list(End   = NA)) %>% 
   #IND Orissa Nuapada
  lbscripts::mutate_when((reg_name == "IND Orissa Nuapada" & Stage == "Planting" & Season == 3),
                         list(End   = NA)) %>% 
  #IND Uttaranchal Naini Tal
  lbscripts::mutate_when((reg_name == "IND Uttaranchal Naini Tal" & Stage == "Planting" & Season == 1),
                         list(End   = NA)) %>% 
  #IND Orissa Malkangiri: Peak == 0. Resete to midpoint S/E
  lbscripts::mutate_when((reg_name == "IND Orissa Malkangiri" & Stage == "Planting" & Season == 2),
                         list(Peak = 258)) %>% 
  #several places with "319   305   334" --> Remove Start
  lbscripts::mutate_when((peakstart_lgt == -14),
                         list(Start = NA)) %>% 
  #several places with "30   -5   35" --> Remove Start
  lbscripts::mutate_when((peakstart_lgt == -5),
                         list(Start   = NA)) %>% 
    # re-create the diffs for test purposes
  dplyr::mutate(endstart_lgt = (End - Start),
                peakstart_lgt = (Peak - Start), 
                endpeak_lgt   = (End - Peak)) 

```
Let's see again the histograms: 

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
print(ggplot(out_atlas) + 
        geom_histogram(aes(x = endstart_lgt), binwidth = 7)  + 
        facet_wrap(~Stage) + theme_bw() +
        ggtitle("Differences between End and Start in riceatlas"))

print(ggplot(out_atlas) + 
        geom_histogram(aes(x = peakstart_lgt), binwidth = 7) + 
        facet_wrap(~Stage) + theme_bw() + 
        ggtitle("Differences between Peak and Start in riceatlas"))

print(ggplot(out_atlas) + 
        geom_histogram(aes(x = endpeak_lgt), binwidth = 7) + 
        facet_wrap(~Stage) + theme_bw() + 
        ggtitle("Differences between End and Peak in riceatlas"))

```

Ok. Now we are in much better shape (note that I never toched the "Peak" date, which 
we take as "reference"). `End - Start` is always positive, as well as `End - Peak`. 
Still some polygons with "anomalously" large ranges, but we can live with those.

Let's see the PHL region:
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
in_phl <- out_atlas %>% 
  dplyr::filter(cty == "PHL" & Region == "Region 3 - Central Luzon")

ggplot(data = in_phl, aes(x = Sub_region)) +
  geom_point(aes(y = Peak, color = Season), position = position_dodge(.3)) + 
  geom_errorbar(aes(ymax = End, ymin = Start, color = Season), 
                position = position_dodge(width = .3)) +
  facet_wrap(~Stage) + coord_flip() + theme_bw() + ylab("DOY") + xlab("RiceAtlas polygon")
```

Almost there. NOw I'll just have to reconcile the planting and the harvesting: 
if negative, subtract 365. (I'll do it later today)
