---
title: "RiceAtlas Comparison"
output:
html_document:
highlight: pygments
theme: journal
toc: yes
toc_depth: 2
editor_options: 
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sprawl)
library(ggplot2)
library(ggjoy)
library(tibble)
library(raster)
library(phenoriceR)
library(dplyr)
library(mapview)
library(RColorBrewer)

mapviewOptions(basemaps = c("Esri.WorldImagery", "OpenStreetMap.DE"),
               raster.palette = colorRampPalette(brewer.pal(9, "RdYlGn")),
               vector.palette = colorRampPalette(brewer.pal(9, "YlGnBu")),
               na.color = "transparent",
               layers.control.pos = "topright")


```

Here I am going to document the passages required to compare PhenoRice results
with RiceAtlas

# Preprocessing

## Select an area

For simlpicity, let's work on a selected area (in this case, the Philippines) 
(for Bhogendra: I created a folder structure containing all required input 
data. It would be good if you could replicate it on your machine so that we
can more easily share the code)

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# This is the main folder containing preprocessed PhenoRice global mosaics
# in the form of multiband TIFF images for each parameter (in "/orig"), as 
# well as the results of "decircularization (in "/decirc")
mosaic_folder  <- "/home/lb/my_data/prasia/Data/orig_mosaic/param_series/"
subsets_folder <- "/home/lb/my_data/prasia/Data/subsets"
make_folder(subsets_folder, type = "dirname", verbose = T)
subset_name <- "PHL"
in_country  <- "PHL"
boundmask   <- sprawl::get_boundaries(in_country, level = 1) %>%
  sf::st_as_sf() %>% 
  sf::st_combine() %>%
  sf::st_sf(id = 1, sf_column_name = ".") 
plot_vect(boundmask, scalebar_dist = 200, show_axis = TRUE)

# Extract PhenoRice data for the PHL subset area - cropped and masked results are 
# placed in "prasia/Data/Subsets/PHL"

pr_extract_subarea(mosaic_folder,
                boundmask,
                subset_name,
                out_folder  = subsets_folder)

#Extract RiceAtlas data for PHL

```

## Visualize data

Extract and visualize rice atlas data for an area and "variable -  in this case
"Peak" planting date. 

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
in_txt <- "/home/lb/my_data/prasia/various/NormalizedDB.txt"
in_riceatlas_shp = "/home/lb/my_data/prasia/Data/vector/Ricetlas/riceatlas_asia.shp"
riceatlas_phl <- read_riceatlas(in_txt, in_riceatlas_shp, "PHL")
peak_plant_s1 <- subset(riceatlas_phl, Pheno_stage == "plant_st1")
plot_vect(peak_plant_s1, fill_var = "Peak", palette_name = "RdYlGn")
```

Here we immediately __see something strange__. The "dark green" polygon has
completely different values. Let's see a histogram:  

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
p <- ggplot(peak_plant_s1, aes(x = doytodate(Peak, 2005))) + 
  geom_histogram(binwidth = 15, fill = "darkgrey", color = "black") +
  scale_x_date(limits = doytodate(c(0,365), 2005) ,
               date_labels = "%b") + 
  theme_bw() 
p

```
Looking at the original table of the `rice_production_v1` shapefile, this is
confirmed (i.e., it is not a data conversion problem). __The "correct" dates for 
the Wet Season of the offending polygons is instead reported in this case in 
"plant_st_2"__:

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
data_leyte <- subset(riceatlas_phl, Sub_region %in% c("Leyte" , "Samar") & 
                       Pheno_stage %in% c("plant_st1", "plant_st2")) %>% 
  select(Sub_region, Season, Peak) 
st_geometry(data_leyte) = NULL
data_leyte$Peak <- doytodate(data_leyte$Peak, year = 2005) %>% 
  format("%d - %b")
knitr::kable(data_leyte)
```

You can see that __the two seasons are "reversed" in Leyte__. The problem is that 
if these kind of errors are "widespread" among the DB, then making the comparisons
with PhenoRice may be difficult. 

Additionally, __comparing the `rice_production_v1` and `rice_calendar` data show 
inconsistencies__. For example, in `rice_calendar`, South Leyte polygon has
correct data, while in `rice_production_v1` they are wrong. However, though 
in general the other polygons have similar values, they are not identical. 

Finally, another problem is that, as in original PhenoRice outputs, __the DOYs are 
not "decircularized"__. This makes things difficult because area with the same 
calendar can have for example "peak date" for planting equal to 360 or 10. 
I will have to think of a way to solve this problem, if we wish to compare 
PhenoRice with RiceAtlas. 

### Update after Bhogendra reprocessing of riceAtlas data

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
in_txt <- "/home/lb/my_data/prasia/various/NormalizedDB_Asia_new.txt"
in_riceatlas_shp = "/home/lb/my_data/prasia/Data/vector/Ricetlas/riceatlas_asia.shp"
riceatlas_phl <- read_riceatlas(in_txt, in_riceatlas_shp, "PHL")
peak_plant_s1 <- subset(riceatlas_phl, Pheno_stage == "plant_st1")
plot_vect(peak_plant_s1, fill_var = "Peak", palette_name = "RdYlGn")
```

``` {r}

riceatlas_phl %>% 
  select(Sub_region, N_seasons, Season, Pheno_stage, Start, Peak, End)

```
Problem is evidently related to not accounting for circular nature of DOY. 

### Further reshaping of riceatlas data

Reshape a bit the dataset: 

```{r reshape_data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Reshape a bit the dataset
in_atlas <- read.csv2(in_txt) %>% 
  # split columns
  tidyr::separate(Pheno_stage, c("Stage", "season"), "_st") %>% 
  # remove empty rows
  dplyr::filter((Start != 0 | Peak != 0 | End != 0)) %>% 
  # create unique name
  dplyr::mutate(reg_name = paste(ISO, Region, Sub_region)) %>% 
  dplyr::mutate(reg_name = stringr::str_replace(reg_name, "NA", "")) %>% 
  # Rename and reorder
  dplyr::select(Objectid, ISO, reg_name, N_seasons, season, 
                Stage, Start, Peak, End, Total_area, Seasonal_area, 
                Seasonal_production, Season, Country, Region, Sub_region) %>% 
  dplyr::rename(cty = ISO, N_seas = N_seasons, Seas_name = Season, Season = season,
                Seas_area = Seasonal_area , Seas_prod = Seasonal_production, 
                Tot_area = Total_area) %>% 
  dplyr::mutate(Stage = factor(Stage, 
                               levels = c("plant", "harv"), 
                               labels = c("Planting","Harvesting"))) %>% 
  tibble::as_tibble()
names(in_atlas)[1] = "OBJECTID"

in_atlas
```

Now let's compute the differences in the Start/Peak/End columns and check their 
distribution: 

```{r}

in_atlas <- in_atlas %>% 
  dplyr::mutate(endstart_lgt  = (End - Start), 
                peakstart_lgt = (Peak - Start),
                endpeak_lgt   = (End - Peak)
  ) %>% 
  dplyr::select(1:9, endstart_lgt, peakstart_lgt, endpeak_lgt, everything())

p <- ggplot(in_atlas) + 
  geom_jitter(aes(x = "", y = endstart_lgt), size = 0.4, alpha = 0.5) + 
  facet_wrap(~Stage) + theme_bw()
p + ggtitle("Distribution of the differences between End and Start in riceatlas")

p <- ggplot(in_atlas) + 
  geom_jitter(aes(x = "", y = peakstart_lgt), size = 0.4, alpha = 0.5) + 
  facet_wrap(~Stage) + theme_bw()
p + ggtitle("Distribution of the differences between Peak and Start in riceatlas")

p <- ggplot(in_atlas) + 
  geom_jitter(aes(x = "", y = endpeak_lgt), size = 0.4, alpha = 0.5) + 
  facet_wrap(~Stage) + theme_bw()
p + ggtitle("Distribution of the differences between End and Peak in riceatlas")
```

From this, it is obvious that the dataset is generally coherent, but also that in some cases the Start/Peak/End values do not account for "circularity". Therefore we can have cases such as:
- `Start = 354 ,Peak = 3, End = 18` --> `End <=Start` and `Peak <= Start` --> leading to negative outliers
- `Start = 4, Peak = 292, End = 34` --> `End <=Start` and `Peak <= Start` --> leading to positive outliers

, plus other less common problems (e.g., `Start = 32 ,Peak = 32, End = 0`,
`Start = 213 ,Peak = 227, End = 213`, `Start = 277, Peak = 19, End = 307`), as well
as obvious mistakes (e.g., `Start = 74, Peak = 74, End = 196`)

I tried to automatically fix, first of all, the more common problems: 

```{r}

out_atlas <- in_atlas %>%
  # add 365 to End/Peak where differences are negative
  lbscripts::mutate_when((endstart_lgt < 0 & endstart_lgt < 0), 
                         list(End = End + 365)) %>%
  # re-create the diffs for test purposes
  dplyr::mutate(endstart_lgt = (End - Start),
                peakstart_lgt = (Peak - Start), 
                endpeak_lgt   = (End - Peak)) %>% 
  lbscripts::mutate_when((endpeak_lgt > 250 & peakstart_lgt < -250), 
                         list(Start = Start - 365, 
                              End   = End - 365)) %>% 
  # re-create the diffs for test purposes
  dplyr::mutate(endstart_lgt = (End - Start),
                peakstart_lgt = (Peak - Start), 
                endpeak_lgt   = (End - Peak)) %>% 
  
  lbscripts::mutate_when(endpeak_lgt > 300, 
                         list(End   = End - 365))%>% 
  # re-create the diffs for test purposes
  dplyr::mutate(endstart_lgt = (End - Start),
                peakstart_lgt = (Peak - Start), 
                endpeak_lgt   = (End - Peak)) 

p <- ggplot(out_atlas) + 
  geom_jitter(aes(x = "", y = endstart_lgt), size = 0.4, alpha = 0.5) + 
  facet_wrap(~Stage) + theme_bw()
p + ggtitle("Distribution of the differences between End and Start in riceatlas")

p <- ggplot(out_atlas) + 
  geom_jitter(aes(x = "", y = peakstart_lgt), size = 0.4, alpha = 0.5) + 
  facet_wrap(~Stage) + theme_bw()
p + ggtitle("Distribution of the differences between Peak and Start in riceatlas")

write.table(out_atlas,
           file = "/home/lb/my_data/prasia/various/NormalizedDB_Asia_temp.txt", 
           sep = "  ",
          row.names = FALSE)

```

Ok. Now we are in much better shape (We can fix the last inconsistencies by
hand). Now let's work on assigning all homogenizing assignment to "Season Number". 
Let's use the PHL example again for testing. 
```{r}

in_txt <- "/home/lb/my_data/prasia/various/NormalizedDB_Asia_temp.txt"
riceatlas_phl <- read_riceatlas(in_txt, in_riceatlas_shp, "PHL") 
riceatlas_phl

```

We first have to reconcile so that on "seasons" spanning the end of the year 
we either always use  `-15, 0, 15`  OR `355, 365, 375`. I decide for the latter
because the negatives can bring in trouble later. Thertefore, I check to see
if there is any negative number in "Start" or "Peak" and add 365 

```{r}

out_riceatlas_phl <- riceatlas_phl %>% 
  lbscripts::mutate_when((Start < 0 & End > 0), 
                         list(Start = Start + 365, 
                              Peak  = Peak + 365), 
                              End = End +365) %>% 
  lbscripts::mutate_when((Peak < 0 & End > 0), 
                         list(Peak = Peak + 365)) 

```


peak_plant_s1 <- subset(riceatlas_phl, Stage == "Planting" & Season == 1)
plot_vect(peak_plant_s1, fill_var = "Peak", palette_name = "RdYlGn")
